---
title: "STAT 547M Homework 06: Data Wrangling"
author: "Rebecca Asiimwe"
date: '2018-11-01'
output: 
  html_document:
      theme: united
      toc: yes
      toc_level: 4
      keep_md: true
      hightlight: tango
      number_sections: false
      toc_float:
        collapse: false
        smooth_scroll: false
  github_document:
    toc: yes
---

#### [Link to Homework-05 homepage](https://github.com/STAT545-UBC-students/hw06-rasiimwe)

### Assignment Goals:
Assignment 6 covers concepts that are key for us to successfully work with non-numeric data like characters and factors. This work covers the following key areas as prior described in the [Readme.md] repo](https://github.com/STAT545-UBC-students/hw06-rasiimwe) file of this homework:

1. Writing R functions
2. Working with regular expressions and character data
3. Working with purrr, list-columns, nested data frames


#### Loading required packages:

```{r results='hide', message=FALSE, warning=FALSE}
suppressPackageStartupMessages(library(stringr)) #avails string functions 
suppressPackageStartupMessages(library(gapminder)) #loading the Gapminder excerpt
suppressPackageStartupMessages(library(ggplot2)) #will be required to make some plots
suppressPackageStartupMessages(library(dplyr))# for required data manipulation
suppressPackageStartupMessages(library(quantreg))  #quantile regression
suppressPackageStartupMessages(library(MASS)) #Functions and datasets to support Venables and Ripley
suppressPackageStartupMessages(library(tidyverse)) #provides system of packages for data manipulation
suppressPackageStartupMessages(library(knitr))
suppressPackageStartupMessages(library(plotly))#for renderin interactive plots
suppressPackageStartupMessages(library(purrr))#to provide tools for working with functions and vectors like map()
suppressPackageStartupMessages(library(RColorBrewer))#provide color palette
suppressPackageStartupMessages(library(tibble))
suppressPackageStartupMessages(library(repurrrsive))
suppressPackageStartupMessages(library(tidytext))
```

## Assignment Activity 2: Writing functions (Part I)

###### Question 2 overview:

###### Write one (or more) functions that do something useful to pieces of the Gapminder or Singer data. It is logical to think about computing on the mini-data frames corresponding to the data for each specific country, location, year, band, album, … This would pair well with the prompt below about working with a nested data frame, as you could apply your function there. Make it something you can’t easily do with built-in functions. Make it something that’s not trivial to do with the simple dplyr verbs. The linear regression function presented [here](http://stat545.com/block012_function-regress-lifeexp-on-year.html) is a good starting point. You could generalize that to do quadratic regression (include a squared term) or use robust regression, using MASS::rlm() or robustbase::lmrob().


### Creating a Function for Regression Model fitting

R provides several regression methods that model the relationship between variables; that is the dependent and independent variables. The methods for robust regression that are provided by R also help us work with data that has outliers with specificity in fitting datasets with large outliers. To kick off this exploration of building functions in this case, one on fitting regression models, I will create a data set to work with.

### Creating a dataframe to work with

```{r}
ctry <- "Kuwait" #assigning Kuwait to variable ctry for country
(data <- gapminder %>% 
  filter(country == ctry)) %>%  #filtering the gapminder dataset for data on Kuwait and passing this to the kable() function for better tabular presentation
  kable()
```

The table above shows the dataframe that has been created for subsequent analysis. This data is an extract of the gapminder datset from which the country `Kuwait` has been filtered.


### Ploting the data
It is always good practice to plot the data to derive a better perspective of the variables being analysed and the relationship between them. 

```{r}
data %>% 
  ggplot(aes(year, gdpPercap)) + 
  geom_point() + 
  geom_smooth(method = "lm", se = TRUE) + #'se' denotes error confidence bounds
  theme_bw() +
  ggtitle("GDP Per Capita of Kuwait over the years") +
  theme( plot.title = element_text(hjust = 0.5)) 
```

The above plot shows a scatter plot of the gdp per capita of Kuwait over the years. We can see that Kuwait's gpd per capita kept reducting over the years! In the code snippet that generated the above code, the `lm` method was used to fit the linear model. 

### Fitting regression models 
#### (linear model, quantile regression and a robust regression model)
To create and fit the various specified regression models, I will begin by getting some code that works:

```{r}
#linear model
liner_model_fit <- lm(gdpPercap ~ year, data=data)  #fitting linear model on gdpPercap ~ year in the data dataset
coef(liner_model_fit) #function to extract model coefficients from the fitted model
```

The above output from the `lm` function is a vector of estimated intercept and slope, from the linear regression of gdp per capita on year. From the above output, we can see that the gdp per capita of Kuwait is estimated to drop given the negative correlation and slope between gdp per capita and year.

### Predicting y intercept given x

```{r}
year_var <- 1993
lm(data$gdpPercap ~ I(data$year - year_var))#here I am not piping ( %>% ) to avoid coercing the formula to a data frame
```

Let's say, I would like to see the y intersept of a certain corresponding year; to do this, I used the `I()` function. In the above example, I used the year `1993` which is not directly shown in the previous plot (GDP Per Capita of Kuwait over the years). Running the above line of code, then helps me know that in 1993, the gdp per capita of Kuwait is estimated to have been  43949.


#### Creating other regression models
Below, I create other regression models as guided by [this tutorial on robust regression](http://www.alastairsanderson.com/R/tutorials/robust-regression-in-R/). Here I use quatile regression to model the median of gdpPercap as a function of year. I also use and iteratively re-weighted least squares fit to down-weight outliers according to how far they are from the best-fit line. This also iteratively re-fits the model until convergence is achieved.
 
```{r}
#Quantile regression:
quantile_regression_fit <- rq(gdpPercap ~ year, data=data)

#Iteratively re-weighted least squares fit: 
least_squares_fit <- rlm(gdpPercap ~ year, data=data)
```


### Plotting the data using various regression methods
Hereunder, I use ggplot to plot the data as well as fit and plot the regression lines, with standard error confidence bounds displayed as a semi-transparent, shaded envelope (se=TRUE).

```{r, message=FALSE}
data %>% 
  ggplot(aes(year, gdpPercap)) +                                  
  geom_point()+                                                   
  geom_smooth(method="lm", aes(colour="liner_model_fit"), se=TRUE) +   #fitting and plotting lm model + envelope       
  geom_smooth(method="rq", aes(colour="quantile_regression_fit"), se=FALSE) + #fitting and plotting quadratic model          
  geom_smooth(method="rlm", aes(colour="least_squares_fit"), se=FALSE) +  #fitting & plotting robust model      
  labs(colour=NULL) +         
  theme_bw() +
  ggtitle("GDP Per Capita of Kuwait over the years") +
  theme( plot.title = element_text(hjust = 0.5))
```

By default, the fitting is applied separately to each specified "group" of data, which is identified in this case by the different colours & line types applied to each unique value of the "type" column in the supplied data frame.


### Turning created working code into a function 
Below is a small function that outputs the y intersept and slope on supplied function arguments

```{r}
model_fit <- function(df, y_var, yr) {
  if (!is.numeric(yr)) {
        stop(paste("Expecting year to be numeric. You supplied", typeof(year)))}
  else {
  fit <- lm(y_var ~ I(year - yr), df)
  setNames(coef(fit), c("intercept", "slope"))
  }
}

year <- 1993
model_fit(data, data$gdpPercap, year) #here y_var is the users choice of the y variable
#if "November" or "2018" is supplied as year we get the following function derived error
#Error in model_fit(data, year) : Expecting year to be numeric. You supplied character
```

From the above function, we see that we still get the same intersept and slope as got in the section on `Predicting y intercept given x`

### Testing my model_fit() function on other arguments
Below I test the model_fit() function on lifeExp instead of gdpPercap as seen earlier:

```{r}
year <- 2000
model_fit(data, data$lifeExp, year)
```

Testing `model_fit()` on lifeExp and year == 2000 from the data dataset outputs the y intersept as `77.4674788` and the slope as `0.4168364`. Next, I will validate that the output seen above can be generated using the independent I() function: 

```{r}
year_var <- 2000
lm(data$lifeExp ~ I(data$year - year_var))
```

We see that we still get similar results as those got from the function `model_fit`. The function works!!! Yaaayyyy..... 💃 💃 💃 

### Testing the 'model_fit()' function on a different data set

```{r}
#creating a dataset as an extract of the gapminder dataset from which the country `Canada` has been filtered.
ctry2 <- "Canada" 
(data2 <- gapminder %>% 
  filter(country == ctry2)) %>% 
  kable()
```

#### Deriving y intercept given year 1993 on data2

```{r}
year2 <- 1993
lm(data2$gdpPercap ~ I(data2$year - year2))
```
#### Using the created function `model_fit()` to accomplish the same (Deriving y intercept given year 1993 on data2)

```{r}
year2 <- 1993
model_fit(data2, data2$gdpPercap, year2)

```

From the above outputs, we see that the y intercept generated by the I() function and that generated by the created function is the same! Nice!!


### Function derived plot on the three regression models

Below, I create a function to allow model fiting on all numeric variables of the gapminder dataset given a certain supplied year. Here, I also allow the user to specify the x and y labels and plot title based on the variables they would like to analyse.

```{r}
#Function arguments
#df         the dataset
#y          y-axis variable
#x          x-axis variable
#xlab       x-axis label
#ylab       y-axis label
#title      plot title

plot_model_fit <- function(df, y, x, xlab, ylab, title){
  
  liner_model_fit <- lm(y ~ x, data=df) 
  coef(liner_model_fit)
  quantile_regression_fit <- rq(y ~ x, data=df)
  least_squares_fit <- rlm(y ~ x, data=df)
  
  plot <- df %>% 
    ggplot(aes(x, y)) +                                 
    geom_point()+                                                   
    geom_smooth(method="lm", aes(colour="liner_model_fit"), se=TRUE) +        
    geom_smooth(method="rq", aes(colour="quantile_regression_fit"), se=FALSE) +           
    geom_smooth(method="rlm", aes(colour="least_squares_fit"), se=FALSE) +        
    labs(colour=NULL) +         
    theme_bw() +
    ggtitle(title) +
    theme( plot.title = element_text(hjust = 0.5))+
    labs(x=xlab, y=ylab)
  return(plot)
}

xlab <- "year"
ylab <- "gdpPercap"
title <- "Function Derived Plot - GDP Per Capita of Kuwait over the Years"
plot_model_fit(data, data$gdpPercap, data$year, xlab, ylab, title)

```

#### Running the `plot_model_fit()` function on lifeExp instead of gdpPercap:

```{r, warning=FALSE}
xlab <- "year"
ylab <- "LifExp"
title <- "Function Derived Plot - LifeExp of Kuwait over the Years"
plot_model_fit(data, data$lifeExp, data$year, xlab, ylab, title)
```

From the above code snippet, I change the 2nd function argument to lifeExp instead of gdpPercap. I also supplied a different plot title and a different y-axis lable to nicely represent the variables being looked into.

## Assignment Activity 2: Writing functions (Part II)

### Creating a function that generates ggplots of my choice

Most of my previous work on data exploration and analysis using R's ggplot2 package has involved heavy usage of repetitive ggplot code. To avoid this repetitive process, I have decided to create a function that creates various plots for me.  Here I will test my function with a few plot types. To kick off this section, I will create code that generates working plots 

### Code that generates a working scatter plot 

```{r}

#defining color palette to be used in subsequent sections
colourCount = length(unique(gapminder$continent))
getPalette = colorRampPalette(brewer.pal(9, "Set1"))

gapminder %>% 
    ggplot(aes(lifeExp, gdpPercap)) +
  geom_point(alpha=0.2)+
  geom_smooth(method=lm)+
    scale_y_log10() +
    labs(x="lifeExp", y="gdpPercap") + 
    theme_bw() +
    ggtitle(paste("Gapminder: gdpPercap Vs lifeExp")) +
    theme(plot.title = element_text(hjust = 0.5))
```


### Code that generates a working box plot

```{r}
gapminder %>% 
  ggplot(aes(reorder(continent, gdpPercap), gdpPercap)) + 
  geom_boxplot(fill=getPalette(colourCount), alpha=0.4) +
  theme_bw() +
  geom_jitter(alpha=0.05) +
  ggtitle("Gapminder: gdpPercap across all continents") +
  theme(plot.title = element_text(hjust = 0.5)) +
  labs(x="")
    
```

### Code that generates a working violin plot

```{r}
gapminder %>% 
  ggplot(aes(reorder(continent, gdpPercap),gdpPercap)) + 
  geom_violin() +
  theme_bw() +
  geom_jitter(alpha=0.05) + 
  labs(x="") +
  ggtitle("Gapminder: gdpPercap across all continents") +
  theme(plot.title = element_text(hjust = 0.5)) 
 
```

### Creating the function to generate plots of choice 
In the code below, I create a function to generate `scatter`, `box` and `violin` plots as example plots. In this function I also specify whether the plots should be rendered as a plotly plot by specifying "TRUE" or "FALSE" for the plot2 function argument. 

```{r}

#initiating funciton "myplot()" with arguments needed to create various plots
# Arguments
#x                    dataset to use for the plots
#type                 plot type to be generated - this could be a scatter plot, violin plot, facets or box plot 
#x_variable           the variable to be used on the x axis
#y_variable           the variable to be used on the y axis
#factor               specified factor to be used in a plot
#jitter               argument to specifies whether a plot should have geom_jitter or not, takes on TRUE or FALSE values
#order                argument that specifies whether a plot should have its factors ordered, takes on TRUE or FALSE values
#xlab                 argument that takes the x-axis lable as defined by the user
#ylab                 argument that takes the y-axis lable as defined by the user
#title                argument that takes the plot title  as defined by the user

myplot <- function(x, type, x_variable, y_variable, factor1, jitter, order, xlab, ylab, title, plot2){
  base <- x %>% 
    ggplot(aes(x_variable, y_variable)) +
    scale_y_log10() +
    labs(x = xlab, y = ylab) + 
    theme_bw() +
    ggtitle(title) +
    theme(plot.title = element_text(hjust = 0.5))
  
  if (type == "scatter"){ 
    if (plot2 == "TRUE"){
    scatter_plot <-  base + 
      geom_point(alpha = 0.2) +
      geom_smooth(method = lm)
    scatter_plot2 <- ggplotly(scatter_plot)
    return(scatter_plot2)}
  
  else if(plot2 == "FALSE"){
    scatter_plot <-  base + 
      geom_point(alpha = 0.2) +
      geom_smooth(method = lm)
    return(scatter_plot)}}
  
  else if (type == "violin"){
    if(plot2 == "TRUE"){
      violin_plot <- x %>% 
        ggplot(aes(reorder(factor1, y_variable),y_variable)) + 
        geom_violin() +
        theme_bw() +
        geom_jitter(alpha = 0.05) +
        ggtitle(title) +
        theme(plot.title = element_text(hjust = 0.5)) +
        labs(x = "", y = ylab)
       violin_plot2 <- ggplotly(violin_plot)
      return(violin_plot2)
      
    }
    else if (plot2 == "FALSE"){
      violin_plot <- x %>% 
        ggplot(aes(reorder(factor1, y_variable),y_variable)) + 
        geom_violin() +
        theme_bw() +
        geom_jitter(alpha=0.05) +
        ggtitle(title) +
        theme(plot.title = element_text(hjust = 0.5)) +
        labs(x = "", y = ylab)
      return(violin_plot)
    }
    else if (type == "box"){
      if(jitter == "TRUE" & plot2 == "TRUE"){
        box_plot <- x %>% 
          ggplot(aes(reorder(factor1, y_variable), y_variable)) + 
          geom_boxplot(fill=getPalette(colourCount), alpha=0.4) +
          theme_bw() +
          geom_jitter(alpha = 0.05) +
          ggtitle(title) +
          theme(plot.title = element_text(hjust = 0.5)) +
          labs(x = "")
        return(box_plot)
        }
      else if (jitter == "TRUE" & plot2=="FALSE"){
        box_plot <- x %>% 
          ggplot(aes(reorder(factor1, y_variable), y_variable)) +
          geom_boxplot(fill = getPalette(colourCount), alpha=0.3) +
          theme_bw() +
          geom_jitter(alpha = 0.1)+
          ggtitle(title) +
          theme(plot.title = element_text(hjust = 0.5)) +
          labs(x = "")
        return(box_plot)
        }
      else if (jitter == "FALSE" & plot2 == "FALSE"){
        box_plot <- x %>% 
          ggplot(aes(reorder(factor1, y_variable),y_variable)) + 
          geom_boxplot(fill=getPalette(colourCount), alpha=0.3) +
          theme_bw() +
          ggtitle(title) +
          theme(plot.title = element_text(hjust = 0.5)) +
          labs(x = "")
        return(box_plot)
      }
    }
    
  }
  else print ("sorry, I can't provide the plot you specified - please check the plot arguments supplied. Plots being tested in this funciton can only be scatter, box or violin plots")
}

```

### Running function `myplot()` to create a scatter plot with "plotly == "TRUE"`

```{r}
#defining my prefered axis lables
xlab <- "lifeExp"
ylab <- "gdpPercap"
title <- "Gapminder: lifeExp Vs gdpPercap (Plotly turned on)"

myplot(gapminder, "scatter", gapminder$lifeExp, gapminder$gdpPercap, gapminder$continent, "TRUE", "TRUE", xlab, ylab, title, "TRUE")

```

### Running function `myplot()` to create a violin plot with `plotly == "FALSE"`

```{r}
xlab <- " "
ylab <- "gdpPercap"
title <- "Gapminder: gdpPercap across all continents (Plotly turned off)"
myplot(gapminder, "violin", gapminder$lifeExp, gapminder$gdpPercap, gapminder$continent, "TRUE", "TRUE", xlab, ylab, title, "FALSE")
```


### Running function `myplot()` to create a box plot with `jitter == "TRUE"` and `plotly == "TRUE"`

```{r}
xlab <- " "
ylab <- "gdpPercap"
title <- "Gapminder: gdpPercap across all continents (Jitter and Plotly turned on)"
myplot(gapminder, "box1", gapminder$lifeExp, gapminder$gdpPercap, gapminder$continent, "TRUE", "TRUE", xlab, ylab, title, "FALSE")
```

In the above code snippet, I provide a plot type that is not known to the function. This throws the function generated  error, "sorry, I can't provide the plot you specified - please check the plot arguments supplied. Plots being tested in this funciton can only be scatter, box or violin plots"


## Assignment Activity 5: Working with lists

###### Question 5 overview:

###### Work through and write up a lesson from the [purrr tutorial](https://jennybc.github.io/purrr-tutorial/index.html): 
- [Trump Android Tweets](https://jennybc.github.io/purrr-tutorial/ls08_trump-tweets.html)
- [Simplifying data from a list of GitHub users](https://jennybc.github.io/purrr-tutorial/ls02_map-extraction-advanced.html)

### Lessons from purrr
This section walks through the [application of purrr](https://jennybc.github.io/purrr-tutorial/index.html) in data wrandling as nicely availed by Jenny Brian and inspired by [David Robinson’s blog post](http://varianceexplained.org/r/trump-tweets/) on "Text analysis of Trump’s tweets confirms he writes only the (angrier) Android half". For this analysis, I will use the same trump_tweets data set and provide various wrangling variants and a cheatsheet of working with purrr. 

### Loading the trump_tweets dataset

```{r}
#loading the trump_tweets dataset into global environment (trump_tweets_df)
load(url("http://varianceexplained.org/files/trump_tweets_df.rda"))
```

### trump_tweets - preliminary data inspection

Next, I will inspect the dataset to help me get a better perspective of it, that is to know what it contains and its dimensions. To achieve this, I will run a few sniffing functions like head(), str(), dim(), colnames() using a function I have created to help me achieve this. My motivation for creating this `inspect() function` is that in all my work with R, I noticed that before I start working on any dataset, I need to do a preliminary inspection of the data. To avoid repetitive writing of the same code that does data inspectsion, below, I create a small function that takes any dataframe and does preliminary sniffing on it. For illustration, I have selected a few exploratory functions.

### Function to automate preliminary data exploration/inspection for any dataset

```{r}

# Arguments
#x      this is the dataset on which prelimnary data sniffing will be done
inspect <- function(x) {
  
    f <- validObject(x, test = FALSE, complete = FALSE) #testing if supplied object is a valid R object
    
    if (f == "FALSE")
      print("x is not a valid R object, check the object name")
    
    else {
     
        classx <- x %>% 
          class
        
        dimx <- x %>% 
          dim
        
        strx <- x %>% 
          str 
        
        #headx <- x %>% 
         # head(5) %>% 
          #knitr::kable() #needs knitr call as oppsed to singularily calling kable()
        
        fx <-  sapply(colnames(x), function(y) class(x[[y]]))
        
        return(list(classx, dimx, strx, fx))
      }
     
}

```

#### Running the created function `inspect()` to inspect the trump_tweets_df dataset.

```{r}
inspect(trump_tweets_df) #running the inspect() function 
```

From the above output we see that the trump_tweets_df is a `tbl_df` inherited from class `tbl` and `data.frame`. We can also see that it has	1512 observations (rows) and  16 variables (columns) prefixed with a `$`. In the created function, I also use the function `sapply()` to help me derive each column name and datatype as this data is helpful when working with `purr` functions like `map()`

### Working with str_detect(), str_split(), str_replace(), substring()

str_detect() is a function used to check whether word or group of words exists in a string or character vector like  sentences we commonly use. First, I will subset the data to a small piece that I can work with.

```{r}
tweets <- trump_tweets_df$text #subseting the data to work with a small piece of it
words <- regex <- "badly|crazy|weak|spent|strong|dumb|joke|guns|funny|dead" #common words associated with Trump tweets
tweets <- tweets[c(2, 4, 6, 8)]
str(tweets) #character of 4 elements
tweets
```
In the above code snippet, we see that the object tweets is of type chatacter and contains 4 elements. To support subsequent analyses, an object `words` of type character has also been created to hold common words/ regular expressions that are associated with trump tweets.

#### Deriving element at index 2 and the number of distinct words therein - using str_split() and str_detect()
```{r}
tweets[2]
```

The above code ouput line shows the second element in tweets. These elements (words) will further be used in following sections. Let us now extract the number of words.

```{r}
str_split(tweets[2], pattern=" ")[[1]] %>%  #split by space " "
  length() #how many words
str_split(tweets[2], pattern=" ") %>% 
  head()
```

In the above code output, we see by the application of `str_split` that the number of words are 23. Let's say, I have these many words or a sentence, and I would like to find a word or words in a string, to a accomplish this, we can apply the `str_detect()` function to find words for example `economic policy` and `Morell` in  tweets[2] as seen in the code below.

```{r}
str_detect(tweets[2], "economic policy") #false

str_detect(tweets[2], "Morell")#true
```

From the above output, if the pattern being searched for is found in a string, R returns `TRUE`, otherwise it will return `FALSE`. We can also replace words as seen in the code snippet below:

#### Working with str_replace()

```{r}
#replace
tweets[2] %>% 
    str_replace("former", replacement = "current")

```

We can see that the word `former` as seen in tweets[2] "Michael Morell, the lightweight former Acting Director of C.I.A., and a man who has made serious bad calls, is a total Clinton flunky!" has been replaced with  `current`. A similar function to str_replace is the `substring()` function that provides replacement and more operations, however the use of substring() dose replacement based on position as shown below.


#### Working with substring() 
The substring function is a powerful R function that can be used either to extract parts of a character string, or to change the values of parts of character strings. For example running `substring("STAT547M is awesome",1,4)` would return `STAT`

#### Replacing values using the substring function:

```{r}
string<-"Trump has funny words!!!"
substring(string,11,16)="weird"
string

substring(string,11,16)="an interesting choice of" #note truncation - call for recycling
string

```
Notice that "funny" is being replaced with "weird" in output line one. This works perfectly because the replacing and replaced elements are of the same length. In output line 2, however, we see that if the replacing element has length > than the element to be replaced, extra chatacters are truncated.


### Working with gregexpr() and map()
The gregexpr() function is a powerful R function that enables retrieval of matching substrings.
At certain points in time, we may need to work with non rectangular data such as those generated by JSON or XML. We can apply Purr's `map()` function to dig information out of non-rectangular data structures and create neat data frames. 
In the following sections, I walk us through the following funcitons: `map()`, `unnest()` and `enframe()`.


```{r}

found <- gregexpr(words, tweets)
typeof(found)
found %>% 
  str(max.level = 1) 

map_int(found, length)              # purrr way of finding lenghts of elements
map_int(found[2], length)

str(found[[1]], list.len = 4)
#listviewer::jsonedit(found)
```

The above output shows us that the variable `found` is of type and class `list` and when run through the `str()` function shows that the variable `found` has 4 elements when searched through tweets. Here each element is an integer vector and shows -1  if no match is found. 


### Putting the above functions together to create a tibble 

```{r}
df2 <- tibble(text = tweets,
       first = gregexpr(regex, tweets)) %>% 
  mutate(match_length = map(first, ~ attr(.x, which = "match.length")),
         last = map2(first, match_length, ~ .x + .y - 1)) %>%
  select(-match_length) %>% 
  pmap(substring)

```

### Visualizing the trump_tweets dataset

```{r}
#creating a tidier dataset

regex_words <- "([^A-Za-z_\\d#@']|'(?![A-Za-z_\\d#@]))" 
tweets2 <- trump_tweets_df %>%
  filter(!str_detect(text, "^RT")) %>%
  mutate(text = str_replace_all(text, "https://t.co/[A-Za-z\\d]+|http://[A-Za-z\\d]+|&amp;|&lt;|&gt;|RT|https", "")) %>%
  unnest_tokens(word, text, token = "regex", pattern = regex_words) %>%
  filter(!word %in% stop_words$word,
         str_detect(word, "[a-z]"))

tweets2 %>%
  count(word, sort=TRUE) %>%
  filter(substr(word, 1, 1) != '#', # omiting hashtags
         substr(word, 1, 1) != '@', # omiting Twitter handles
         n > 80) %>% # only most common words
  mutate(word = reorder(word, n)) %>%
  ggplot(aes(word, n)) +
  geom_bar(stat = 'identity', fill=c("gray50")) +
  xlab(NULL) +
  ylab(paste('(Word count)', sep = '')) +
  ggtitle(paste('common words in tweets')) +
  theme(legend.position="none") +
  coord_flip() +
  theme_bw() +
  theme(plot.title = element_text(hjust = 0.5))
```


### Working with the gh_users dataset for a more interesting view on chatacter data

This dataset is a list with 6 components, each representing a GitHub user. Each user's component is a list of length 30, containing information such as username, GitHub id as will be shown below.

```{r}
#inspect(gh_users)
str(gh_users, max.level = 1)
str(gh_users[[1]], list.len = 6)
#listviewer::jsonedit(gh_users)
```

From the above output, we see that gh_users is a list with 6 main elements `login`, `id`, `avatar_url`, `gravatar_id`, `url` and `html_url`. We can also see that it is not easy to map various elements together. To do this, we can use the `map()` function to reach into each  list (a user's list) and map an element of interest to another element say username. This is shown in the code snippet below. We can also test other data types with map, for example map_lgl() for logical values, map_int() for integers and map_dbl() for doubles. In the below example, I show a mapping of names to login ids.

```{r}

map(gh_users, "login")
map(gh_users, 2)

map_chr(gh_users, "login")#type-specific variant of map() - chr for character
map_chr(gh_users, 2)

```

### Extracting multiple values

```{r}
gh_users[[6]][c("name", "login", "id", "location")]

```
In instances where we need to extract multiple elements we can use square bracket indexing and the character vector of the names of our desired elements as shown the example output above. Here we see an extract of the name, login, id and location of user in index position 6(Maëlle Salmon). We can also extracting multiple elements per user by mapping [ function as seen below.

```{r}
x <- map(gh_users, magrittr::extract, c("login", "name", "id", "location"))
str(x[3:4])
```

### Working with the map_df() function.

When working with non rectangular data, we can use the `map_df()` function to help us create neat data frames. A case in point is the example code shown below in which we use `map_df()`to create a data frame of one row per user and `login`, `id`, `avatar_url`, `gravatar_id`, `url` and `html_url` as variables. 

```{r}
map_dataframe <- map_df(gh_users, `[`, c("id", "login", "name", "url"))

map_dataframe %>% 
  kable()

```

The above table is the neat data frame created using `map_df()`. Let us look into the data types of each of the variables in the created dataframe.

```{r}
fx <-  sapply(colnames(map_dataframe), function(y) class(map_dataframe[[y]]))
fx
```

From the above ouput, we can see that the variable `id` is an integer, `login`, `name` and `url` are characters. 


### Creating a list inside a data frame

We can also encapsulate a list inside a dataframe!! and in the example in the code snipped below, we aim at creating a data frame with one row per github repository identified by user name. To achieve this, we create a variable and pass the github repos list to it including user names. 

```{r}
(names <- map_chr(gh_repos, c(1, 4, 1)))
```

After parsing the user names to variable names, we can now create a data frame as shown in the code snippet below:

```{r}
(repo_df <- gh_repos %>%
    set_names(names) %>% 
    enframe("username", "gh_repos"))
```

#### Smell testing the created dataframe

```{r}
repo_df %>% 
  class
```

```{r}
repo_df %>% 
  dim
```


```{r}
sapply(colnames(repo_df), function(y) class(repo_df[[y]]))
```

From the above output we see that the created data frame is of class `tbl_df` inherited from classes `tbl` and `data.frame` respectively. The tibble has 6 observations (rows) and 2 variables (colums): username and gh_repos which are of type character and list respectively.


### Sources to acknowledge:

[Notes on regression](http://www.alastairsanderson.com/R/tutorials/robust-regression-in-R/)

[Jenny Brian's tutorial on regression](http://stat545.com/block012_function-regress-lifeexp-on-year.html)

[purrr tutorial](https://jennybc.github.io/purrr-tutorial/index.html): 

[Trump Android Tweets](https://jennybc.github.io/purrr-tutorial/ls08_trump-tweets.html)

[Simplifying data from a list of GitHub users](https://jennybc.github.io/purrr-tutorial/ls02_map-extraction-advanced.html)

[More on Trump Tweets](https://pushpullfork.com/mining-twitter-data-tidy-text-tags/)



